Gompertz growth as infinitely correlated non-local growth

### 2023-09-20 arxiv2/new_draft.tex
add levitt and zondas paper

Finally, on the topic of disease modelling, the Gompertz model has been shown by Wang et al. \cite{wang2012richards} to be excluded from a communicable disease paradigm where only local disease spread is possible. They used the common Susceptible-Infected-Recovered (SIR) model by \citet{kermack1927contribution} augmented by the Richards parameter and showed that the allowed parameter range with the communicable disease assumption does not cover the parameter value taken in the Gompertz scenario. Given this observation by Wang et al., any mortality pattern that shows a Gompertz characteristic cannot be a result of a communicable disease, but rather from a ubiquitous stressor. While many studies show the SIR model (and its extensions) as the fitted model in epidemics \cite{carletti2020covid,cooper2020sir,postnikov2020estimation,munoz2021sir,cooper2022dynamical,saikia2021covid}, some recent studies have successfully fitted a Gompertz model \cite{Ohnishi2020,Rypdal2020,Catala2020,rodrigues2020monte,Levitt2020}.

@@

For growth within biological organisms, it has been well established that the mediating field is the electromagnetic field as discussed by R.O. Becker (see \citet{becker1984electromagnetic} and references within), which suggests that the accompanying growth or decay pattern in these processes follows the Gompertz model as the result of a ubiquitous activating field. For growth or decay in larger ecological or societal systems other catalysts can be air pollution, oxygen deprivation, or even instant psychologically induced catalysts.

@@

It is worth noting that the original formulation of the Gompertz model made by Gompertz himself was concerning mortality rates as a function of age. He noticed that mortality rates increased at a geometric rate with age. In this context the ubiquitous stressor postulated above would be the living conditions on our planet, for which time acts as a proxy. One could further postulate this stressor as a field surrounding the earth, e.g. in its electromagnetic footprint often related with the Schumann Resonance \cite{cherry2002schumann,schumann1952strahlungslosen}. It is indeed curious that we do not see mortality rates behave as linearly dependent on time, for which a Gaussian curve would be appropriate. Nor do we see any finite-ordered polynomial time dependence on mortality rates. Instead we see growth of mortality rates as sourced from other entities at all interaction orders, which in light of our discussion above is equivalent to mortality coming from an environmental stressor.

>>>>

On the topic of disease modelling, the Gompertz model has been shown by Wang et al. \cite{wang2012richards} to be excluded from a subset of models in the communicable disease paradigm. They used the common Susceptible-Infected-Recovered (SIR) model by \citet{kermack1927contribution} augmented by the Richards parameter and showed that the allowed parameter range with the communicable disease assumption does not cover the parameter value taken in the Gompertz scenario. If one incorporates a social proximity network on which the disease spreads, \citet{zonta2022virus} observed through computer simulations that a near Gompertz growth can occur but only for scale-free networks, a type of network that has been shown to be rare in human social networks \cite{clauset2009power,broido2019scale,holme2019rare}. Given these observations, any mortality pattern that shows a Gompertz characteristic seems not be a result of a communicable disease, but rather from a ubiquitous stressor. While many studies show the SIR model (and its extensions) as the fitted model in epidemics \cite{carletti2020covid,cooper2020sir,postnikov2020estimation,munoz2021sir,cooper2022dynamical,saikia2021covid}, some recent studies have successfully fitted a Gompertz model \cite{Ohnishi2020,Rypdal2020,Catala2020,rodrigues2020monte,Levitt2020}.

@@

For many growth processes within biological organisms, it has been well established that the mediator is the electromagnetic field \cite{becker1984electromagnetic}, which suggests that the accompanying growth or decay pattern in these processes follows the Gompertz model as the result of a ubiquitous activating field. For growth or decay in larger ecological or societal systems other catalysts can be air pollution, oxygen deprivation, or even instant psychologically induced catalysts.

@@

We finally note that the original formulation of the Gompertz model made by Gompertz himself was concerning mortality rates as a function of age. He noticed that mortality rates increased at a geometric rate with age. In this context the ubiquitous stressor postulated above would be the living conditions on our planet, for which time acts as a proxy. One could further postulate this stressor as a field surrounding the earth, e.g. in its electromagnetic footprint often related with the Schumann Resonance \cite{cherry2002schumann,schumann1952strahlungslosen}. It is indeed curious that we do not see mortality rates behave as linearly dependent on time, for which a Gaussian curve would be appropriate. Nor do we see any finite-ordered polynomial time dependence on mortality rates. Instead we see growth of mortality rates as sourced from other entities at all interaction orders, which in light of our discussion above is equivalent to mortality coming from an environmental stressor.
